---
title: "Project 2"
author: "Robert West"
date: "6/30/2020"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(knitr)
```


## Set up the call to Analysis.RMD
```{r}
days = c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday")
out_file = paste0(days, ".md")

params = lapply(days, FUN=function(x){
  return(list(day=x))
})

reports = tibble(days, out_file, params)
reports
```


## Generating the reports
```{r, eval=F}
library(rmarkdown)

apply(reports, MARGIN = 1, FUN=function(x){
  render(input = "Analysis.Rmd", output_file = x[[2]], params=x[[3]])
})
```


Here are the the reports that are generated for each day:   

* The analysis for [Monday is available here](Monday.md).    
* The analysis for [Tuesday is available here](tuesday.md).    
* The analysis for [Wednesday is available here](wednesday.md).    
* The analysis for [Thursday is available here](thursday.md).    
* The analysis for [Friday is available here](friday.md).    
* The analysis for [Saturday is available here](saturday.md).    
* The analysis for [Sunday is available here](sunday.md).    










